Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-01 doc/thesis/appendix.Rnw
moved objective stuff to methods and expanded

\chapter{Mathematical models, likelihood, and Bayesian inference}
\label{chp:prelim}

@@

\chapter{GLM results}

<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
    format.glm <- function (df) {
        df[Parameter == "alpha", Parameter := "$\\alpha$"]
        df[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
        df[,Estimate := paste0("$", sapply(Estimate, pp, eq=FALSE, thres=NA), "$")]
        df[,`Standard error` := paste0("$", sapply(`Standard error`, pp, eq=FALSE, thres=NA), "$")]
        setnames(df, "p", "$p$-value")
        df
    }

    alpha.glm <- format.glm(alpha.glm)
    I.glm <- format.glm(I.glm)
    #m.glm <- format.glm(m.glm)
    #N.glm <- format.glm(N.glm)

    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
                        "$\\alpha$ to true values of BA parameters.")
    capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
                  "inverse link function. Coefficients are interpretable as",
                  "additive effects on the inverse of the mean error.")
    print(xtable(alpha.glm, label="tab:glmalpha", caption=c(capt, capt.short)),
          include.rownames=FALSE, sanitize.text.function=identity,
          sanitize.colnames.function=identity)

    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
                        "$I$ to true values of BA parameters.")
    capt <- paste(capt.short, "GLM was fitted with a gaussian distribution and",
                  "inverse link function. Coefficients are interpretable as",
                  "additive effects on the inverse of the mean error.")
    print(xtable(I.glm, label="tab:glmI", caption=c(capt, capt.short)),
          include.rownames=FALSE, sanitize.text.function=identity,
          sanitize.colnames.function=identity)

    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
    #                    "$m$ to true values of BA parameters.")
    #capt <- paste(capt.short, "GLM was fitted with a Poisson distribution and",
    #              "log link function. Coefficients are interpretable as",
    #              "additive effects on the exponential of the mean error.")
    #print(xtable(m.glm, label="tab:glmm", caption=c(capt, capt.short)),
    #      include.rownames=FALSE, sanitize.text.function=identity,
    #      sanitize.colnames.function=identity)

    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
    #                    "$N$ to true values of BA parameters.")
    #capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
    #              "inverse link function. Coefficients are interpretable as",
    #              "additive effects on the inverse of the mean error.")
    #print(xtable(N.glm, label="tab:glmN", caption=c(capt, capt.short)),
    #      include.rownames=FALSE, sanitize.text.function=identity,
    #      sanitize.colnames.function=identity)
@

>>>>

\chapter{Mathematical models, likelihood, and Bayesian inference}
\label[appendix]{chp:prelim}

@@

%\chapter{GLM results}
%<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
%    format.glm <- function (df) {
%        df[Parameter == "alpha", Parameter := "$\\alpha$"]
%        df[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
%        df[,Estimate := paste0("$", sapply(Estimate, pp, eq=FALSE, thres=NA), "$")]
%        df[,`Standard error` := paste0("$", sapply(`Standard error`, pp, eq=FALSE, thres=NA), "$")]
%        setnames(df, "p", "$p$-value")
%        df
%    }
%
%    alpha.glm <- format.glm(alpha.glm)
%    I.glm <- format.glm(I.glm)
%    #m.glm <- format.glm(m.glm)
%    #N.glm <- format.glm(N.glm)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$\\alpha$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(alpha.glm, label="tab:glmalpha", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$I$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a gaussian distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(I.glm, label="tab:glmI", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%    #                    "$m$ to true values of BA parameters.")
%    #capt <- paste(capt.short, "GLM was fitted with a Poisson distribution and",
%    #              "log link function. Coefficients are interpretable as",
%    #              "additive effects on the exponential of the mean error.")
%    #print(xtable(m.glm, label="tab:glmm", caption=c(capt, capt.short)),
%    #      include.rownames=FALSE, sanitize.text.function=identity,
%    #      sanitize.colnames.function=identity)
%
%    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%    #                    "$N$ to true values of BA parameters.")
%    #capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
%    #              "inverse link function. Coefficients are interpretable as",
%    #              "additive effects on the inverse of the mean error.")
%    #print(xtable(N.glm, label="tab:glmN", caption=c(capt, capt.short)),
%    #      include.rownames=FALSE, sanitize.text.function=identity,
%    #      sanitize.colnames.function=identity)
%@

### 2016-06-01 doc/thesis/appendix.tex
moved objective stuff to methods and expanded

@@

%\chapter{GLM results}
%
%<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
%    format.glm <- function (df) {
%        df[Parameter == "alpha", Parameter := "$\\alpha$"]
%        df[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
%        setnames(df, "p", "$p$-value")
%        df
%    }
%
%    alpha.glm <- format.glm(alpha.glm)
%    I.glm <- format.glm(I.glm)
%    m.glm <- format.glm(m.glm)
%    N.glm <- format.glm(N.glm)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$\\alpha$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Gamma distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(alpha.glm, label="tab:glmalpha", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$I$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Gamma distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(I.glm, label="tab:glmI", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$m$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Poisson distribution and",
%                  "identity link function. Coefficients are interpretable as",
%                  "additive effects on the exponential of the mean error.")
%    print(xtable(m.glm, label="tab:glmm", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$N$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
%                  "identity link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(N.glm, label="tab:glmN", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%@

>>>>

\chapter{Mathematical models, likelihood, and Bayesian inference}
\label[appendix]{chp:prelim}

A \defn{mathematical model} is a formal description of a hypothesized
relationship between some observed data, $x$ and outcomes $y$. A
\defn{parametric} model defines a family of possible relationships between data
and outcomes, parameterized by one or more numeric parameters $\theta$. A
\defn{statistical} model describes the relationship between data and outcomes
in terms of probabilities. Statistical models define, either explicitly or
implicitly, the probability of observing $y$ given $x$ and, if the model is
parametric, $\theta$. Note that it is entirely possible to have no data $x$,
only observed outcomes $y$. In this case, a model would describe the process by
which $y$ is generated.

To illustrate these concepts, consider the well-known linear model. For
clarity, we will restrict our attention to the case of one-dimensional data and
outcomes where $x = \set{x_1, \ldots, x_n}$ and $y = \set{y_1, \ldots, y_n}$
are vectors of real numbers. The linear model postulates that the outcomes are
linearly related to the data, modulo some noise introduced by measurement
error, environmental fluctuations, and other external factors. Formally, $y_i =
\beta x_i + \varepsilon_i$, where $\beta$ is the slope of the linear
relationship, and $\varepsilon_i$ is the error associated with measurement $i$.
We can make this model a statistical one by hypothesizing a distribution for
the error terms $\varepsilon_i$; most commonly, it is assumed that they are
normally distributed with variance $\sigma$. In mathematical terms, $y_i \sim
\beta x_i + \N(0, \sigma^2)$, where ``$\sim$'' means ``is distributed as''. We
can see from this formulation that the model is parametric, with parameters
$\theta$ = ($\beta$, $\sigma$). Moreover, we can write down the probability
density $\pi$ of observing outcome $y_i$ given the parameters,
\[
  \pi(y \mid \beta, \sigma) = 
  \prod_{i=1}^n f_{\N(0, \sigma^2)} (y_i - \beta x_i),
\]
where $f_{\N(0, \sigma^2)}$ is the probability density of the normal
distribution with mean zero and variance $\sigma^2$. Note that we are treating
the $x_i$ as fixed quantities and therefore have not conditioned the
probability density on $x$. Also, we have assumed that all the $y_i$ are
independent.

For a general model, the probability density of $y$ given the parameters
$\theta$ is also known as the \defn{likelihood}, written $\L$, of $\theta$.
That is, $\L(\theta \mid y) = f(y \mid \theta)$ for the model's \gls{pdf} $f$.
The higher the value of the likelihood, the more likely the observations $y$
are under the model. Thus, the likelihood provides a natural criterion for
fitting the model parameters: we want to pick $\theta$ such that the
probability density of our observed outcomes $y$ is as high as possible. The
parameters that optimize the likelihood are known as the \textit{\gls{ML}}
estimates, denoted $\hat{\theta}$. That is,
\[
  \hat{\theta} = \argmax_\theta\; \L(\theta \mid y).
\]
\Gls{ML} estimation is usually performed with numerical optimization. In the
simplest terms, many possible values for $\theta$ are examined, $\L(\theta \mid
y)$ is calculated for each, and the parameters that produce the highest value
are accepted. Many sophisticated numerical optimization methods exist, although
they may not be guaranteed to find the true \gls{ML} estimates if the
likelihood function is multi-modal.

\Gls{ML} estimation makes use only of the data and outcomes to estimate the
model parameters $\theta$. However, it is frequently the case that the
investigator has some additional information or belief about what $\theta$ are
likely to be. For example, in the linear regression case, the instrument used
to measure the outcomes may have a well-known margin of error, or the sign of
the slope may be obvious from previous experiments. The Bayesian approach to
model fitting makes use of this information by codifying the investigator's
beliefs as a \defn{prior distribution} on the parameters, denoted
$\pi(\theta)$. Instead of considering only the likelihood, Bayesian inference
focuses on the product of the likelihood and the prior, $f(y \mid \theta)
\pi(\theta)$. Bayes' theorem tells us that this product is related to the
\textit{posterior distribution} on $\theta$,
\begin{align}
  f(\theta \mid y) 
    = \frac{f(y \mid \theta) \pi(\theta)}
           {\int f(y \mid \theta) \pi(\theta) \d \theta}.
  \label{eq:bayes}
\end{align}
In principle, $f(y \mid \theta) \pi(\theta)$ can be optimized numerically just
like $\L(\theta \mid y)$, which would also optimize the posterior distribution.
The resulting optimal parameters are called the \gls{MAP} estimates. However,
from a Bayesian perspective, $\theta$ is not a fixed quantity to be estimated,
but rather a random variable with an associated distribution (the posterior).
Therefore, the \gls{MAP} estimate by itself is of limited value without
associated statistics about the posterior distribution, such as the mean or
credible intervals. Unfortunately, to calculate such statistics, it is
necessary to evaluate the normalizing constant in the denominator of
\cref{eq:bayes}, which is almost always an intractable integral.

A popular method for circumventing the normalizing constant is the use of
\gls{MCMC} to obtain a sample from the posterior distribution. \Gls{MCMC} works
by defining a Markov chain {\color{red}\sout{whose states are indexed by
possible model parameters. The transition probability from state $\theta_1$ to
state $\theta_2$ is taken to be}} {\color{blue}\uline{on the space of possible
model parameters. The transition density from parameters $\theta_1$ to
$\theta_2$ is taken to be}}
\[
  \min\left(1, \frac{f(y \mid \theta_2) \pi(\theta_2) q(\theta_2, \theta_1)}
                    {f(y \mid \theta_1) \pi(\theta_2) q(\theta_1, \theta_2)} \right),
\]
where $q(\theta, \theta')$ is a symmetric \defn{proposal distribution} used in
the algorithm to generate the chain. The stationary distribution of this Markov
chain is equal to the posterior distribution on $\theta$. Therefore, if a long
enough random walk is performed on the chain, the distribution of states
visited will be a Monte Carlo approximation of $f(\theta \mid y)$, from
which we can calculate statistics of interest. Actually performing this random
walk is straightforward and can be accomplished via the Metropolis-Hastings
algorithm~\autocite{metropolis1953equation,hastings1970monte} (\cref{alg:mh}).

\begin{algorithm}
  \caption{Metropolis-Hastings algorithm for Markov chain Monte Carlo.}
  \begin{algorithmic}
    \State Draw $\theta$ according to the prior $\pi(\theta)$
    \Loop
      \State Propose $\theta'$ according to $q(\theta, \theta')$
      \State Accept $\theta \gets \theta'$ with probability
      $\min \left( 1, 
       \dfrac{f(y \mid \theta') \pi(\theta') q(\theta', \theta)}
             {f(y \mid \theta\phantom{'}) \pi(\theta\phantom{'}) q(\theta, \theta')}
       \right)$
    \EndLoop
  \end{algorithmic}
  \label{alg:mh}
\end{algorithm}

@@

%\chapter{GLM results}
%<<glm_tables, echo=FALSE, warning=FALSE, error=FALSE, results="asis">>=
%    format.glm <- function (df) {
%        df[Parameter == "alpha", Parameter := "$\\alpha$"]
%        df[,p := paste0("$", sapply(p, pp, eq=FALSE), "$")]
%        df[,Estimate := paste0("$", sapply(Estimate, pp, eq=FALSE, thres=NA), "$")]
%        df[,`Standard error` := paste0("$", sapply(`Standard error`, pp, eq=FALSE, thres=NA), "$")]
%        setnames(df, "p", "$p$-value")
%        df
%    }
%
%    alpha.glm <- format.glm(alpha.glm)
%    I.glm <- format.glm(I.glm)
%    #m.glm <- format.glm(m.glm)
%    #N.glm <- format.glm(N.glm)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$\\alpha$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(alpha.glm, label="tab:glmalpha", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%                        "$I$ to true values of BA parameters.")
%    capt <- paste(capt.short, "GLM was fitted with a gaussian distribution and",
%                  "inverse link function. Coefficients are interpretable as",
%                  "additive effects on the inverse of the mean error.")
%    print(xtable(I.glm, label="tab:glmI", caption=c(capt, capt.short)),
%          include.rownames=FALSE, sanitize.text.function=identity,
%          sanitize.colnames.function=identity)
%
%    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%    #                    "$m$ to true values of BA parameters.")
%    #capt <- paste(capt.short, "GLM was fitted with a Poisson distribution and",
%    #              "log link function. Coefficients are interpretable as",
%    #              "additive effects on the exponential of the mean error.")
%    #print(xtable(m.glm, label="tab:glmm", caption=c(capt, capt.short)),
%    #      include.rownames=FALSE, sanitize.text.function=identity,
%    #      sanitize.colnames.function=identity)
%
%    #capt.short <- paste("Parameters of a fitted GLM relating error in estimated",
%    #                    "$N$ to true values of BA parameters.")
%    #capt <- paste(capt.short, "GLM was fitted with a Gaussian distribution and",
%    #              "inverse link function. Coefficients are interpretable as",
%    #              "additive effects on the inverse of the mean error.")
%    #print(xtable(N.glm, label="tab:glmN", caption=c(capt, capt.short)),
%    #      include.rownames=FALSE, sanitize.text.function=identity,
%    #      sanitize.colnames.function=identity)
%@

### 2016-06-01 doc/thesis/intro.tex
moved objective stuff to methods and expanded

\subsection{Transmission trees}

@@

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the \emph{posterior distribution}
\begin{align}
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)}.
  \label{eq:post}
\end{align}
\uline{As outlined in} \cref{sec:obj}, \uline{both the likelihood $\Pr(T \mid
\theta)$ and the normalizing constant $\Pr(T)$ are likely computationally
intractable. Hence, rather than computing the posterior distribution
analytically, we will approximate it using a \defn{Monte Carlo} approach.
The fundamental idea behind Monte Carlo methods is succinctly expressed by
\textcite{liu2001theoretical}}:
\begin{quote}
    \uline{Monte Carlo's view of the world is that any probability distribution $\pi$, 
    regardless of its complexity, can always be \emph{represented} by a
    discrete sample from it. By ``represented'', we mean that any computation
    of expectations using $\pi$ can be replaced to an acceptable degree of
    accuracy by using the empirical distribution resulting from the discrete
    sample.}
\end{quote}
\uline{In other words, if we are able to sample enough points from a
distribution of interest, we will be able to make reasonably accurate
statements about the distribution itself. For example, the expected value of
the distribution can be estimated by the sample's population mean. The reason
Monte Carlo methods will be useful in this work is that algorithms exist for
obtaining samples from distributions, including as the aforementioned
posterior, that are analytically intractable and from which direct sampling is
not possible } (for a review see \autocite{robert2004monte}).
\uline{\Gls{SMC}}~\autocite{doucet2000sequential, doucet2001introduction,
liu2008monte} \uline{is one such algorithm.}

\uline{ \Gls{SMC} is also known as the \emph{particle filter}. Rather than
sampling points one at a time from the target distribution, \gls{SMC} considers
a population of points or ``particles'', here denoted $\set{x^{(k)}}$. The
particles are associated with weights, $\set{w(x^{(k)})}$. These weighted
particles are a representation for the target distribution. For example, the 
expected value of the distribution is approximated by}
\[
  \frac{1}{n} \sum_{k=1}^n x^{(k)} w(x^{(k)}),
\]
\uline{where $n$ is the number of particles. Initially, the particles do not
represent the target distribution but rather a more tractable distribution from
which direct sampling is straightforward. The word ``sequential'' is used to
describe the iterative process of perturbation, resampling, and reweighting
applied to the particles in such a way that they converge, collectively, to a
sample from the target. }

\uline{In this work, the distribution of interest is the posterior distribution
\ref{eq:post}. The particles are particular values of the parameters $\theta$
of the contact network model being studied. If we were taking a typical
Bayesian Monte Carlo approach to this problem, the particles would end up
weighted by their posterior probability and distributed in such a way that the
weighted population was a reasonable representation of $\Pr(\theta \mid T)$.
This turns out to be not straightforward in our case, as we will address in
}\cref{sec:abc}. \uline{For now, nothing is lost by \ldots}}

@@

{\color{blue}\uline{ Our method integrates the four distinct research areas
  just described: phylogenetics, contact networks, \acrlong{SMC}, and
  \acrlong{ABC}. The first two topics together form the problem domain.
  Phylogenetic data is the input to our method, while estimates of the
  parameters of contact network models are the desired output. The latter two
  topics define the algorithm and statistical framework that our inference
method will use. }

\uline{Having introduced transmission trees and contact network models, we are
now in a position to argue that approximate Bayesian computation is an
appropriate tool to apply towards our main objective: fitting network models to
phylogenetic data. Let $M$ be a contact network model having parameters
$\theta$.

}}

>>>>

\subsection{Transmission trees}
\label{subsec:tt}

@@

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the \emph{posterior distribution}}
\begin{align}
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)}.
  \label{eq:post}
\end{align}
\uline{As outlined in} \cref{sec:obj}, \uline{both the likelihood $\Pr(T \mid
\theta)$ and the normalizing constant $\Pr(T)$ are likely computationally
intractable. Hence, rather than computing the posterior distribution
analytically, we will approximate it using a \defn{Monte Carlo} approach.
The fundamental idea behind Monte Carlo methods is succinctly expressed by
\textcite{liu2001theoretical}}:
\begin{quote}
    \uline{Monte Carlo's view of the world is that any probability distribution $\pi$, 
    regardless of its complexity, can always be \emph{represented} by a
    discrete sample from it. By ``represented'', we mean that any computation
    of expectations using $\pi$ can be replaced to an acceptable degree of
    accuracy by using the empirical distribution resulting from the discrete
    sample.}
\end{quote}
\uline{In other words, if we are able to sample enough points from a
distribution of interest, we will be able to make reasonably accurate
statements about the distribution itself. For example, the expected value of
the distribution can be estimated by the sample's population mean. The reason
Monte Carlo methods will be useful in this work is that algorithms exist for
obtaining samples from distributions, including as the aforementioned
posterior, that are analytically intractable and from which direct sampling is
not possible } (for a review see \autocite{robert2004monte}).
\uline{\Gls{SMC}}~\autocite{doucet2000sequential, doucet2001introduction,
liu2008monte} \uline{is one such algorithm.}

\uline{\Gls{SMC} is also known as the \emph{particle filter}. Rather than
sampling points one at a time from the target distribution, \gls{SMC} considers
a population of points or ``particles'', here denoted $\set{x^{(k)}}$. The
particles are associated with weights, $\set{w(x^{(k)})}$. These weighted
particles are a representation for the target distribution. For example, the 
expected value of the distribution is approximated by}
\[
  \frac{1}{n} \sum_{k=1}^n x^{(k)} w(x^{(k)}),
\]
\uline{where $n$ is the number of particles. Initially, the particles do not
represent the target distribution but rather a more tractable distribution from
which direct sampling is straightforward. The word ``sequential'' is used to
describe the iterative process of perturbation, resampling, and reweighting
applied to the particles in such a way that they converge, collectively, to a
sample from the target.}

\uline{In this work, the distribution of interest is the posterior distribution}
\ref{eq:post}. \uline{The particles are particular values of the parameters
$\theta$ of the contact network model being studied. If we were taking a
typical Bayesian Monte Carlo approach to this problem, the particles would end
up weighted by their posterior probability and distributed in such a way that
the weighted population was a reasonable representation of $\Pr(\theta \mid
T)$. This turns out to be not straightforward in our case, as we will address
in} \cref{sec:abc}. \uline{For now, nothing is lost by \ldots}}

@@

{\color{blue}\uline{ Our method integrates the four distinct research areas
  just described: phylogenetics, contact networks, \acrlong{SMC}, and
  \acrlong{ABC}. The first two topics together form the problem domain.
  Phylogenetic data is the input to our method, while estimates of the
  parameters of contact network models are the desired output. The latter two
  topics define the algorithm and statistical framework that our inference
method will use. }}
